---
title: "Bonus Question 1"
author: "Ian Murphy"
date: '2019-09-19'
output: 
  html_document:
    pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Bonus Question 1

Load Packages

```{r load packages, warning = FALSE, message = FALSE}
# load your packages here:
library(gapminder)
library(tidyverse)
```

We are given the following code:

```{r}
gapminder %>%
  filter(country == c("Rwanda", "Afghanistan"))
```


And when we run it, we only get a subset of the data from Rwanda and Afghanistan. In order to do this properly, there are 2 different ways which could work. Firstly, we could replace the "==" with "%in%". This will produce the countries of either Rwanda or Afghanistan. Secondly, we could use the "|" operator. These 2 methods are demonstrated below.

```{r}
gapminder %>% 
  filter(country %in% c("Rwanda", "Afghanistan"))

gapminder %>% 
  filter(country == "Rwanda" | country == "Afghanistan")
```

Either of these methods will produce the desired result. 


