---
title: "Assignment 2 Question 1"
author: "Ian Murphy"
date: '2019-09-18'
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 2 - Question 1
## 1.1

Let's import the gapminder dataset, dplyr, and tidyverse.
```{r}
library(gapminder)
library(tidyverse)
suppressPackageStartupMessages(library(tidyverse))
```

Let's use the filter() function to extract 3 countries in the 1970s. The three countries I will use are Canada, France, and Germany.

```{r}
new_gap <- gapminder %>% 
            filter(country == "Canada" | 
                  country == "France" | 
                  country == "Germany") %>% 
            filter(1969 < year & year <= 1980)
```

As you can see, we've extracted 3 countries during the 1970s, called "reduced_gap".

## 1.2

Let's use pipe operator to select country and gdpPercap from the "reduced_gap" in 1.1.

```{r}
new_gap %>% 
  select(country, gdpPercap)
```

## 1.3

Now we will filter gapminder to all entries that have experienced a drop in life expectancy. 

## 1.4 

We will filter gapminder so that it only shows the max GDP Per Capita of each country. We will group_by country, then we will make use of the slice() function, which will only keep the max value.

```{r}
gapminder %>% 
  select(country, gdpPercap) %>% 
  group_by(country) %>% 
  slice(which.max(gdpPercap))
```
This produces the correct table.

## 1.5 

Produce a scatterplot of Canada's life expectancy vs. GDP per Cap, using ggplot2. That is to say, gdpPercap is on x axis, and lifeExp is on y axis. This shows us that as gdpPercap increases, so does Canadian lifeExp.

```{r}
gapminder %>% 
  filter(country == "Canada") %>%
  ggplot(aes(log(gdpPercap), lifeExp)) + geom_point(colour="blue")
```

